### 2020-03-04 manuscript/bef_stability/manuscript.Rmd
Inverse PCA axis

@@

```{r}
myload(temporal_community_metrics, dir = mypath("data"))
myload(synchrony, dir = mypath("data"))
st_basin <- get_basin_station(sf_obj = FALSE)

myload(temporal_network_metrics, dir = mypath("data", "classes"))
temporal_network_metrics_classes <- temporal_network_metrics
myload(temporal_network_metrics, dir = mypath("data", "species"))

stab_div_troph <- temporal_network_metrics_classes %>%
  dplyr::select(station, troph_group) %>%
  unnest(troph_group) %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group),
    log_rich = log10(richness_med),
    log_rich_tot = log10(richness_tot),
    log_bm = log10(biomass_med),
    log_stab = log10(biomass_stab)) %>%
  dplyr::select(station, troph_group, biomass_stab, richness_med, richness_tot, log_rich, log_rich_tot, log_stab, log_bm, biomass_med, richness_tot) %>%
  dplyr::left_join(st_basin, by = "station")
```

@@

```{r}
myload(
  hab_analysis,
  temporal_station_desc,
  temporal_press_polluants,
  geo_station,
  dir = mypath("data")
)
habitat_pressure <- build_habitat_pressure_dataset(
  .habitat_analysis = hab_analysis,
  .tmp_press = temporal_press_polluants,
  .tmp_st_desc = temporal_station_desc,
  .geo_st = geo_station 
)
```

```{r, fig.show = "hide"}
evt_var_mean <- c("flow_med", "temperature_med", "width_river_mean",
  "avg_depth_station_mean", "slope", "alt", "d_source", "strahler", "DBO_med")
evt_var_cv <- c("flow_cv", "temperature_cv", "width_river_cv",
  "avg_depth_station_cv", "DBO_cv")
pca_pressure <- habitat_pressure[, names(habitat_pressure) %in% c(evt_var_mean, evt_var_cv, "station")] %>%
  na.omit

@@

```{r get-data, cache = FALSE}
library(nlme)
library(piecewiseSEM)

myload(nmds, habitat_pressure, dir = mypath("report"))
myload(op_analysis, dir = mypath("data"))

# Compute stability, network metrics, etc with ;he new datasets
com_data <- compute_community_temporal_analysis(.op = op_analysis, dest_dir = mypath("data", "classes"))

# Compute sem dataset 
sem_data <- compute_sem_dataset(
  com = com_data[["tps_com"]],
  network = com_data[["tps_net"]],
  hab_press = habitat_pressure,
  sync = com_data[["sync"]],
  nmds = NULL,
  basin = st_basin
  )
```

@@

library(ggraph)
p_stab_sem_rich <- ggpiecewisesem(stab_sem_rich)

library(DiagrammeR)
graph <- build_diag_from_sem(stab_sem_rich, p_val_thd = 0.05)

p_bm_sem_rich <- ggpiecewisesem(bm_sem_rich)
```

@@

coef_bm_sem <- coef(stab_bm_rich) %>%
  select(- 9) %>%
  filter(P.Value < 0.05)
```

@@

```{r sem-plot, fig.dim = c(10,10)}
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple
to_plot <- c(1, 6, 7)
plot_grid(
  plot_grid(plotlist = map(p_l_rot[to_plot], ~.x$rotated),
      ncol = length(to_plot), labels = LETTERS[1]),
  plot_grid(p_stab_sem_rich, p_bm_sem_rich, labels = LETTERS[2:3]),
  ncol = 1,
  rel_heights = c(.4, .6)
)
```

- Habitat: (+ / -)
  - RC 1: width & depth & flow med, d source, strahler 	/
  - RC 2: T°C med 					/ slope, alt 
  - RC 3: 						/ flow CV, DBO med
  - RC 4: river width med, depth med 			/
  - RC 5: DBO CV 					/

- Stability:
  - Stability component:
    - CV is the main driver of stability:
  - Community structure: 
    - Species richness had no direct effects on CV but on synchrony!  
    - Evenness of biomass dist increased CV but decreased synchrony
    - Trophic level decreased CV
    - Betadiv increases CV
  - Habitat/Env:
    - few direct effect on stability
      - direct: CV_sp dec with increasing RC3, cad enrichment and flow_cv decrease.  
      - direct: CV_sp inc with increasing RC2, cad temperature_med.  
    - indirectly affect stability by modulating community structure,
      particularly richness

- Total biomass:
  - Community structure:
    - Richness increased total biomass / had the bigger effect
    - Even community deceased total biomass
    - Species turnover decreased total biomass  
    - Connectance increased biomass
  - Habitat/Env:
    - very weak to no direct effect on total biomass 
    - Effect of environment is thus mediated by the community structure

@@

The effect of richness on biomass stability follows a power law with a negative
coefficient, meaning that it's a decelerating function.

- R diagramm:
  - https://rstudio-pubs-static.s3.amazonaws.com/78926_5aa94ae32fae49f3a384ce885744ef4a.html

- Latex:
  - csv:
    - https://tex.stackexchange.com/questions/178904/use-datatool-to-read-a-row-from-a-csv-file-then-use-the-variables-in-the-docume
    - https://tex.stackexchange.com/questions/40901/datatool-getting-a-specific-value-given-the-value-of-another-column
  - path: 
- Reporting:
  - equations: https://stats.stackexchange.com/questions/169115/reporting-the-actual-formula-equation-of-an-lme-model-with-factors-used-in-r
  - log-log results:
    - https://dev.to/rokaandy/logarithmic-transformation-in-linear-regression-models-why-when-3a7c
    - https://stats.stackexchange.com/questions/18480/interpretation-of-log-transformed-predictor-and-or-response

@@

@@

>>>>

Cf blutghen for CVsp as portfolio and statistical averaging of species.
Synchrony as insurance hypothesis facing environmental variations.

@@

```{r}
myload(nmds, habitat_pressure, dir = mypath("report"))
st_basin <- get_basin_station(sf_obj = FALSE)
myload(op_analysis, op_analysis_wo_holes, dir = mypath("data"))
myload(biomass_ts_sax, dir = mypath("data"))
op_analysis_bbb <- filter(op_analysis, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)
op_analysis_wo_holes_bbb <- filter(op_analysis_wo_holes, station %in%
  biomass_ts_sax[biomass_ts_sax$sax == "bbb",]$station)

# Compute stability, network metrics, etc with ;he new datasets
com_data <- compute_community_temporal_analysis(.op = op_analysis_bbb)
# Compute sem dataset 
sem_data <- compute_sem_dataset(
  com = com_data[["tps_com"]],
  network = com_data[["tps_net"]],
  hab_press = habitat_pressure,
  sync = com_data[["sync"]],
  nmds = NULL,
  basin = st_basin
  )

stab_div_troph <- com_data$tps_bm_troph %>%
  filter(troph_group != 1) %>%
  mutate(troph_group = as.factor(troph_group),
    log_rich = log10(richness_med),
    log_rich_tot = log10(richness_tot),
    log_bm = log10(biomass_med),
    log_stab = log10(biomass_stab)) %>%
  dplyr::select(station, troph_group, biomass_stab, richness_med, richness_tot, log_rich, log_rich_tot, log_stab, log_bm, biomass_med, richness_tot) %>%
  dplyr::left_join(st_basin, by = "station")
```

@@

```{r}
myload(
  hab_analysis,
  temporal_station_desc,
  temporal_press_polluants,
  geo_station,
  dir = mypath("data")
)
habitat_press <- build_habitat_pressure_dataset(
  .habitat_analysis = hab_analysis,
  .tmp_press = temporal_press_polluants,
  .tmp_st_desc = temporal_station_desc,
  .geo_st = geo_station 
)
```

```{r, fig.show = "hide"}
evt_var_mean <- c("flow_med", "temperature_med", "width_river_mean",
  "avg_depth_station_mean", "slope", "alt", "d_source", "strahler", "DBO_med")
evt_var_cv <- c("flow_cv", "temperature_cv", "width_river_cv",
  "avg_depth_station_cv", "DBO_cv")
pca_pressure <- habitat_press[, names(habitat_press) %in% c(evt_var_mean, evt_var_cv, "station")] %>%
  na.omit

@@

```{r get-data, cache = FALSE}
library(nlme)
library(piecewiseSEM)
```

@@

library(ggraph)
p_stab_sem_rich <- ggpiecewisesem(stab_sem_rich)
p_bm_sem_rich <- ggpiecewisesem(bm_sem_rich)

library(DiagrammeR)

p_format <- "png"
p_width <- 700
p_height <- 700

stab_sem_path <- mypath("manuscript", "bef_stability",
  paste0("stab_sem.", p_format))
bm_sem_path <- mypath("manuscript", "bef_stability",
  paste0("bm_sem.", p_format))

graph_stab_sem <- build_diag_from_sem(stab_sem_rich, p_val_thd = 0.05)
render_graph(graph_stab_sem)
export_graph(graph = graph_stab_sem, file_type = p_format,
  file_name = stab_sem_path,
  height = p_height - p_height*20/100,
  width = p_width
)
#source(mypath("R", "plot_methods.R")) 
graph_bm_sem <- build_diag_from_sem(bm_sem_rich, p_val_thd = 0.05)
render_graph(graph_bm_sem)
export_graph(graph = graph_bm_sem, file_type = p_format,
  file_name = bm_sem_path,
  #file_name = "stab_sem.svg",
  height = p_height,
  width = p_width 
)

library(magick)
library(rsvg)
img_stab_sem <- stab_sem_path %>%
  image_read()
p_stab_sem_rich <- ggdraw() +
  draw_image(img_stab_sem)
img_bm_sem <- bm_sem_path %>%
  image_read()
p_bm_sem_rich <- ggdraw() +
  draw_image(img_bm_sem)

@@

coef_bm_sem <- coef(bm_sem_rich) %>%
  select(- 9) %>%
  filter(P.Value < 0.05)
```

```{r stab-network}
sem_data %>%
  select(sync, cv_sp, ct, t_lvl) %>%
  gather(network, metrics, ct, t_lvl) %>%
  gather(stab_comp, value, sync, cv_sp) %>%
  ggplot(aes(y = value, x = metrics)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(stab_comp ~ network, scales = "free_x")

stab_ct_mod <- lmer(data = sem_data,
  #fixed = log_bm ~ log_rich_tot * troph_group,
  formula = log_stab ~ ct + (1 | basin)
  #random = ~ 1 + log_rich_tot * troph_group | basin,
  #control = ctrl 
)
stab_ct_boot <- bootMer(stab_ct_mod, mySumm2, nsim = 100)
stab_ct_ci <- confint(stab_ct_boot)

stab_ct_sum <- piecewiseSEM::rsquared(stab_ct_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

pred_stab_ct_mod <- ggpredict(stab_ct_mod, terms = c("ct"), type = "fe")
#plot(pred_troph_mod, add.data = TRUE)
stab_ct_pred <- pred_stab_ct_mod %>%
  as_tibble() %>%
  rename(
    log_bm = predicted,
    log_rich_tot = x,
    troph_group = group
  )

# get coeff
coef_stab_ct <- map(list(stab = stab_ct_mod), fixed.effects)

stab_tlvl_mod <- lmer(data = sem_data,
  #fixed = log_bm ~ log_rich_tot * troph_group,
  formula = log_stab ~ t_lvl + (1 | basin)
  #random = ~ 1 + log_rich_tot * troph_group | basin,
  #control = ctrl 
)
stab_tlvl_boot <- bootMer(stab_tlvl_mod, mySumm2, nsim = 100)
stab_tlvl_ci <- confint(stab_tlvl_boot)

stab_tlvl_sum <- piecewiseSEM::rsquared(stab_tlvl_mod, method = "nagelkerke") %>%
  mutate_at(vars(Conditional, Marginal), ~round(., 2))

pred_stab_tlvl_mod <- ggpredict(stab_tlvl_mod, terms = c("t_lvl"), type = "fe")
#plot(pred_troph_mod, add.data = TRUE)
stab_tlvl_pred <- pred_stab_tlvl_mod %>%
  as_tibble() %>%
  rename(
    log_bm = predicted,
    log_rich_tot = x,
    troph_group = group
  )

# get coeff
coef_stab_tlvl <- map(list(stab = stab_tlvl_mod), fixed.effects)

@@

```{r sem-plot, fig.dim = c(10,10)}
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple
to_plot <- c(1, 6, 7)

plot_grid(
  plot_grid(plotlist = map(p_l_rot[to_plot], ~.x$rotated),
      ncol = length(to_plot), labels = LETTERS[1]),
  plot_grid(p_stab_sem_rich, p_bm_sem_rich, labels = LETTERS[2:3], rel_widths = c(.6, .4)),
  ncol = 1,
  rel_heights = c(.4, .6)
)
```

```{r}
total_effect_bm_sem
total_effect_stab_sem
```

---
# Total biomass
---
We found that the primary drivers of total biomass were the total species
richness (TSR) and the average trophic level (Tlvl) ($r_\delta = 0.35$ and
$r_\delta = 0.40$ resp.) whereas connectance had no significant effect  (Fig.
\@ref{fig:sem-plot} C). The stream size had a direct positive effect on total
biomass ($r_\delta = 0.22$) whereas CV of flow and avg DBO increased total
biomass ($r_\delta = - 0.2$). Avg temperature, CV of river size and CV of DBO
had no direct effects on total biomass.

TSR had also a indirect positive effect on total biomass through Tlvl ($r_\delta
= 0.21$), that is, TSR had a total positive effect of $0.56$ on total biomass.
The stream size had a total effect on total biomass as strong as TSR ($r_\delta
= 0.54$), indirectly through Tlvl ($r_\delta = 0.10$) and TSR ($r_\delta = 0.22$).
The avg temperature of the river had an indirect positive effect on total biomass
($r_\delta = 0.48$), through Tlvl ($r_\delta = 0.11$) and TSR ($r_\delta = 0.33$).
Avg DBO, CV DBO and CV of river size had no or weak indirect effect on total
biomass.

---
# Biomass stability
---
\overline{CV_{sp}} ($r_\delta = -1.26$) decreased biomass stability more than
synchrony ($r_\delta = -0.55$, Figure \@ref{fig:sem-plot}). TSR had a strong
positive effect on \overline{CV_{sp}} and a strong negative effect on synchrony
whereas Tlvl and connectance had no significant effects on either
\overline{CV_{sp}} nor synchrony. The avg river size had a small positive direct
effect on synchrony ($r_\delta = 0.19$). The other environmental variables had
no significant direct effects on \overline{CV_{sp}} nor synchrony.

Indirectly through synchrony and \overline{CV_{sp}}, TSR had a total negative
effect on biomass stability ($r_\delta = -0.26$). Avg river size and avg
temperature had also a total effect which was negative on biomass stability
(resp. $r_\delta = -0.24$ and $r_\delta = -0.16$), going through richness 
(resp. $r_\delta = -0.10$ and $r_\delta = -0.16$).

@@

@@

We tested for the first time the determinants of temporal biomass stability in a
foodweb with an empirical dataset at such high temporal and spatial scales.

Surprisingly, we found a negative relationship biomass stability and species
richness whereas there was a positive one between total biomass and species
richness. Those relationships followed a power law, meaning that they represent
respectively accelerating and decelerating functions. Overall, it contrasts
with studies in grasslands [@tilman_biodiversity_2006, @zhang_scale_nodate],
mesocosms [@pennekamp_biodiversity_2018] or other animals
[@olivier_independent_nodate] where positive relationships are typically found
between biomass stability and species richness. Even @franssen_annual_2011 found
that species rich communities are more stable than species poor in streams.

@grace_integrative_2016 found that biomass negatively linked to species richness
in grasslands, notably because of light competition. In freshwater ecosystems,
where resources are based on trophic relationship, TOCOMPLETE.

The total biomass of high and low trophic group trophic groups responded both
positively to species richness but not biomass stability TOREPHRASE. There is a
paradox where the high trophic group is overall more stable than the low but
species richness had a destabilizing effect on the former whereas a stabilizing
on the latter. @shanafelt_david_w._stability_nodate showed in a theoretical
study in trophic chains, the top trophic level is the more stable TOEXPLORE.
The comparison between our study and theoretical studies stays limited because
fishes in our system have a large diet and consume food item of several trophic
level while theoretical studies consider system in which one species of one
trophic level feeds only on the trophic level directly inferior
[@shanafelt_david_w._stability_nodate] ADD REF.

---
# Effect of synchrony and cv_sp  
---
It was unexpected that the average population CV had a stronger negative effect
on biomass stability than synchrony [@bluthgen_land_2016;
@olivier_independent_nodate] *cf ref in bluthgen*. We see two hypothesis that
can raise this effect. First, the opportunity for diet differentiation can be
low in streams compare to other organisms. Bird species are for example quite
specialized in seed, flying or soil arthropods or mammals which gives
opportunities for independent dynamics. Plants also can achieve independent
dynamics either spatial segregation, niche differentiation. Because fish
species even piscivorous one are quite opportunistic in
their foraging. Secondly, in trophic network, amplitude of species biomass
involved in trophic interactions could covary positively TOTHINKABOUT.

---
# Effect of trophic level and connectance are richness effects:
---
Our path analysis reveled that, although species richness has a strong effect on
connectance and the average trophic level, network structure had itself little
to no effect either on synchrony nor average population stability. This result
contradicts precedent theoretical studies on the relationship
between stability and network structure [@angelis_stability_1975, @may_will_1972], although
they considered different stability metrics (i.e. local stability of equilibrium
point). The reason might be that in our inferred network, network structure is
strongly constraints by assembly processes. Indeed, connectance decreased with
increasing species richness [@thebault_stability_2010], because a small fraction
of links are realized because of food diet constraints. In theoretical studies,
connectance can be tuned from 0 (no links) to 1 (all possible links are
present). SAMEFORTLVL
Finally, the biomass stability of freshwater ecosystems can be derived by
species richness because, through trophic structure constraints, species
richness determine largely network structure.

@@

```{r save-plot}

save_plot("~/Documents/thesis/talks/fishcom_fig/stab_div.pdf",
  p_stab_div,
  base_height = 3,
  bg = "transparent",
  ncol = 1)

save_plot("~/Documents/thesis/talks/fishcom_fig/stab_div_troph.pdf",
  p_stab_div_troph + theme(legend.position = "bottom"),
  base_height = 3,
  bg = "transparent",
  ncol = 1)

### 2020-03-04 report/06-pressure.Rmd
Inverse PCA axis

```{r}
myload(temporal_network_metrics, dir = mypath("data", "classes"))
temporal_network_metrics_classes <- temporal_network_metrics
myload(temporal_network_metrics, dir = dest_dir)
myload(community_metrics, temporal_community_metrics, synchrony,
temporal_press_polluants, press_metrics, op_analysis,
  dir = data_common)
```

>>>>

```{r}
myload(temporal_network_metrics, dir = mypath("data", "classes"))
temporal_network_metrics_classes <- temporal_network_metrics
myload(temporal_network_metrics, dir = dest_dir)
myload(community_metrics, temporal_community_metrics, synchrony,
temporal_press_polluants, press_metrics, op_analysis,
  dir = data_common)

myload(yearly_press_interp, dir = mypath())
```

### 2020-03-04 report/08-infer-interactions.Rmd
Inverse PCA axis

# Sanitize opcod:
map_dbl(list(fish_abun, covariates), ~length(unique(.x$opcod))) 
fish_ok <- fish_abun %>%
  filter(opcod %in% unique(covariates$opcod)) %>%
  arrange(opcod)
cov_mat <- covariates %>%
  filter(opcod %in% unique(fish_abun$opcod)) %>%
  arrange(opcod) %>%
  as.matrix
rownames(cov_mat) <- cov_mat[, "opcod"] 
cov_mat  <- cov_mat[, !colnames(cov_mat) %in% "opcod"]

# remove opcod and build abundance matrix
mat_abun <- fish_abun %>%
  spread(sp_class, nind)%>%
  #select(-opcod) %>%
  as.matrix
rownames(mat_abun) <- mat_abun[, "opcod"] 
mat_abun <- mat_abun[, !colnames(mat_abun) %in% "opcod"]

>>>>

# Sanitize opcod:
map_dbl(list(fish_abun, covariates), ~length(unique(.x$opcod))) 
fish_ok <- fish_abun %>%
  filter(opcod %in% unique(covariates$opcod)) %>%
  arrange(opcod)
cov_mat <- covariates %>%
  filter(opcod %in% unique(fish_abun$opcod)) %>%
  arrange(opcod) %>%
  as.data.frame

rownames(cov_mat) <- cov_mat[, "opcod"] 
cov_mat  <- cov_mat[, !colnames(cov_mat) %in% "opcod"]

# remove opcod and build abundance matrix
mat_abun <- fish_abun %>%
  spread(sp_class, nind) %>%
  as.data.frame

rownames(mat_abun) <- mat_abun[, "opcod"] 
mat_abun <- mat_abun[, !colnames(mat_abun) %in% "opcod"]

save_dir <- mypath("report", "stephane")
if (!dir.exists(save_dir)) {
  dir.create(save_dir)
}
mysave(mat_abun, cov_mat, dir = save_dir)
```
